---
title: "Week 08 Assignment"
author: "Max Wagner"
date: "October 18, 2015"
output: html_document
---
***

Loading the two libraries I'll be using, and then loading the delay data csv file.
```{r, message=FALSE}
library(tidyr)
library(dplyr)
delays <- read.csv(file = "https://raw.githubusercontent.com/maxwagner/607/master/week06/delays.csv", header = FALSE)
```


From here, the goal is create a "tidy data" set that is easy to look at and manipulate. The idea would be to have a column for airline, city, on-time/delay, and number of delayed flights. It might be easier to split the table into two tables. One for alaska airlines, and one for am west. Then combine them at the end.

The first step is selecting the data from the alaska section that I want, making the city names the column names, need to unlist and make them chars. Then gathering the data to flip it around to the way I want it to look. Then add in the airline name in a column, and the ontime/delay info. Then rearrange the columns to make more sense.
```{r, warning=FALSE}
delays_ak <- delays[1:3,3:7]
colnames(delays_ak) <- as.character(unlist(delays_ak[1,]))
delays_ak <- delays_ak[-1,]
delays_ak <- gather(delays_ak, city, n)
delays_ak$airline <- "alaska"
delays_ak$delay <- c("on_time", "delayed")
delays_ak <- delays_ak[c("airline", "city", "delay", "n")]
```

Do the same thing for the am west airline.
```{r, warning=FALSE}
delays_aw <- delays[c(1,5,6),3:7]
colnames(delays_aw) <- as.character(unlist(delays_aw[1,]))
delays_aw <- delays_aw[-1,]
delays_aw <- gather(delays_aw, city, n)
delays_aw$airline <- "am west"
delays_aw$delay <- c("on_time", "delayed")
delays_aw <- delays_aw[c("airline", "city", "delay", "n")]
```

Now combine the two tables into one larger one.
```{r}
delays <- rbind(delays_ak, delays_aw)
```

For the final part, I want to use spread to remove the double instances of each city on the delay column.
```{r}
delays <- spread(delays, delay, n)
```

There were a few other options to transform this data, but I thought this way was very simple compared to the other ways. There is nothing overly confusing and gets the job done. The flaw of this method is manually inputting the airline name and the "on_time / delayed" text, but I think the simplicity in other aspects outweighs this drawback. 

A final comparison between the new format and the old format, I think there is a big improvement.
```{r}
delays
read.csv(file = "https://raw.githubusercontent.com/maxwagner/607/master/week06/delays.csv", header = FALSE)
```