---
title: "Week 08 Assignment"
author: "Max Wagner"
date: "October 18, 2015"
output:
  html_document:
    theme: flatly
    highlight: espresso
---
***

Parse the HTML and XML data with the "XML" package. Remove the null junk from the html table names with "stringr."
```{r, message=FALSE}
library(XML)
books_xml <- xmlToDataFrame("http://pastebin.com/raw.php?i=9sq70BWK")
books_html <- data.frame(readHTMLTable("http://pastebin.com/raw.php?i=M1zhB6XB"))
```

Parse the JSON data with "jsonlite" and "plyr" using a rbind fill.
```{r, message=FALSE}
library(jsonlite)
library(plyr)
books_json <- fromJSON("http://pastebin.com/raw.php?i=6UARcHeJ")
books_json <- do.call("rbind.fill", lapply(books_json, as.data.frame))
```

Let's take a look at the 3 data frames for differences...
```{r}
head(books_xml)
head(books_html)
head(books_json)
```

The only difference between the three formats in a data frame form are the column names. This can be fixed with a few quick "stringr" calls.
```{r}
library(stringr)
colnames(books_html) <- str_replace(colnames(books_html), "NULL.", "")
colnames(books_html) <- str_replace(colnames(books_html), "HTML", "html")
colnames(books_json) <- str_replace(colnames(books_json), "book.", "")
```

Let's look again... and now they are identical.
```{r}
head(books_xml)
head(books_html)
head(books_json)
```