---
title: "Project 01 - Chess"
author: "Max Wagner"
date: "September 26, 2015"
output: html_document
---
***

The first few steps include scanning in the text file, and then concatenating it with a " " as a seperator. It might be excessive, but I then extracted each person as a seperate line instead of the 3 they take up in the original file. This makes pulling data simpler later on.
```{r}
library(stringr)
chess <- scan(file = "https://raw.githubusercontent.com/maxwagner/607/master/project01/tournamentinfo.txt", what = "character", skip = "3")
chess <- str_c(chess, collapse = " ")
chess_sep <- unlist(str_extract_all(chess, "\\d+\\s[|]\\s(\\w+[ -]+)+[|]\\d[.]\\d\\s+[|](\\w\\s+\\d*[|]){7}\\s\\w+\\s[|]\\s\\d+\\s[/]\\sR:\\s[:alnum:]+\\s*->\\s*[:alnum:]+\\s[|].+?[|](\\w*\\s[|]){7}"))
```

The next section is pulling each piece that I will eventually neeed. The opponents section can surely be done in a cleaner way, but this gets the job done. The last line is to turn empties into NA's.
```{r}
name <- str_trim(unlist(str_extract(chess_sep, "\\s(\\w+[ -]+)+")))
state <- unlist(str_extract(chess_sep, "[|]\\s\\w{2}\\s[|]"))
state <- str_trim(str_sub(state, 2, -2));
points <- unlist(str_extract(chess_sep, "\\d[.]\\d"))
starting_score <- unlist(str_extract(chess_sep, "R:\\s\\d+"))
starting_score <- substring(starting_score, 4);
opponents <- unlist(str_extract(chess_sep, "[|](\\w\\s+\\d*[|]){7}"))
opponents <- unlist(str_replace_all(opponents, "[:alpha:]+", ""))
opponents <- unlist(str_replace_all(opponents, "[:space:]+", ""))
opponents <- str_sub(opponents, 2, -2)
opponents <- unlist(str_split(opponents, "[|]"))
opponents <- unlist(str_extract(opponents, "\\d+"))
```

The following loops through the opponents id's and grabs the score from the points vector.
```{r}
for (i in 1:length(opponents)) {
  if (!is.na(opponents[i])) {
    opponents[i] <- starting_score[as.numeric(opponents[i])]
  }
}
```

The next step is to convert the scores into a 64 x 7 shape, and then take the mean of each row.
```{r}
opponents <- data.frame(matrix(as.numeric(opponents), nrow = 64, byrow = T))
opponents <- round(apply(opponents, 1, mean, na.rm = T), 2)
```

Lastly, I combine all the need variables into one frame, and write it to a csv file.
```{r}
chess_final <- cbind(name, state, points, starting_score, opponents)
write.csv(chess_final, file = "chess_output.csv")
```

***
A small sample from the top of the output file:
```{r}
head(read.csv("https://raw.githubusercontent.com/maxwagner/607/master/project01/chess_output.csv"))
```

***
I was looking into the ELO scoring part of it a little, and wrote something that would get the estimated outcome, as well as the new scores, when given the player's old score, opponent's score, the outcome of the game, and a k value. Not really useful without pulling a bit more data from the text file.
```{r}
calc_new <- function(old, opp, score, k) {
  ea <- (1 / (1 + 10 ^ ((opp - old) / 400)))
  rounded <- round((old + (k * (1 - ea))), digits = 0)
  return(rounded)
}
```